Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-03 trunk/doc/thesis/report.tex
Only chapter 3 missing

\documentclass[oneside,12pt,a4paper,final]{book}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\section{IPv6 over USB Tunnel}
\paragraph{}
This is a program that is installed in the gateway server machine. As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology. That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
\paragraph{}
The installation and initialization of this component will be given more details in appendix A.
\section{Gateway Packet Transformation Process}
\paragraph{}
This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world. It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
\paragraph{}

\chapter{Findings}
\section{Experiment}
%\subsection{Methodology}
\section{Data Gathering}
\section{Interpretation of Results}

@@

@@

\appendix
\chapter{How to install the system}

\backmatter

\singlespacing
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

>>>>

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\section{IPv6 over USB Tunnel}
\paragraph{}
This is a program that is installed in the gateway server machine. As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology. That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
\paragraph{}
The installation and initialization of this component will be given more details in appendix A.
\section{Gateway Packet Transformation Process}
\paragraph{}
This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world. It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
\paragraph{}
The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
\paragraph{}
To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work. The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel. The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
\paragraph{}
The program has three subcomponents as seen in section 4.8. The three components are the following:
\begin{itemize}
\item Packet sniffer
\item Packet transformer
\item Address translator
\end{itemize}
The following subsections will discuss each of these subcomponent separately from the implementation point of view.

\subsection{Packet sniffer}
\paragraph{}
The packet sniffer is the one in charge of sniffing packet from both network interface card, as Raw Frames. The code snippet shows the functions called to sniff the packets. We make use of the raw socket library that enables us to sniff data link layer frames before being dropped by Netfilter.
The main reason of dropping the packets by Netfilter is that the packets will not addressed to the computer and therefore the computer make send ICMP messages to the sender saying that the route is not the best and therefore the sender will never send packets again which is something we need to avoid to have.

\begin{lstlisting}
sockfd = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL));
if (sockfd < 0)
	exit(1);

while(1)
{
	data_size = recvfrom(sockfd,buf,10000,0
		,(struct sockaddr *)&cliaddr,&clilen);
}

\end{lstlisting}
\paragraph{}
The first line of code is the initialization of the raw socket, we define the type, scope and name space of the packets to be sniffed. This will return a file descriptor that will allow us to call the socket. Afterwards, we start reading packets in an infinite loop.

\paragraph{}
After sniffing a packet, we must check if it is an \gls{ipv4} or \gls{ipv6} packet. The problem is that when reading from the Wi-Fi interface card, we get the whole frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the raw \gls{ipv6} packet. Once a packet is read by the frame, from which interface card came is a problem and therefore a partial solution is found. We cast the packet to Wi-Fi frame and whenever we check for the network protocol, we get two solutions, 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to drop the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
\paragraph{}
In each of those conditions we only select the TCP datagrams and this is why we check if the next header in the IP structure is equal to 6. After that we move the pointer to drop the IP header and only keep the TCP datagram that will have to be sent in a the newly transformed packet. Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.

\begin{lstlisting}
if(header->h_proto == 0) 
{
	if(ip6_hdr->ip6_nxt == 6)
	{
			buf += IP6_HDRLEN;
			... 			
			sendIPV4Packet(..)
	}
}
else if(header->h_proto == 8)
{
	if(ip_hdr->protocol == 6)
	{
		buf += (sizeof(struct ethhdr)+ IP4_HDRLEN);		
		...
		sendIPV6Packet(...)
	}
}
\end{lstlisting}

\subsection{Packet transformer}
\paragraph{}
The packet tranformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets. As stated in section 4.8, all devices in this system have an \gls{ipv4} and \gls{ipv6} address where one of them is not recognized only at the gateway. As the \gls{wsn} excelusively supports \gls{ipv6} and they cannot be addressed without using their \gls{ipv6} address and any other device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} address at once the packets reach the gateway, it transform so that the data carried will be delivered and the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
\paragraph{}
This subcomponent is based on two important functions:
\begin{itemize}
\item sendIPv6Packet(..)
\item sendIPv4Packet(..)
\end{itemize}
\paragraph{}
The idea is the same for except that each technology has its own details such as filling different fields and computing the checksum. Another difference is when creating the \gls{ipv4} one should also create the Wi-Fi frame otherwise it won't be able to be sent through the corresponding network interface card whereas when building \gls{ipv6} packet one should just send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
The function receives the data to be sent which is in all cases a \gls{tcp} datagram along with the size of the datagram, the source and destination addresses that are already translated. Afterwards, the transformator creates the packet appends the datagram, fills the source and destination address and create a socket attached to the corresponding interface card and send the packet.
\begin{lstlisting}
int sendIPV6Packet(char* buffer,int size,char *src ,char *dest)
{
	//CREATE PACKET WITH ALL ITS DETAILS
	...
	memcpy (frame + IP6_HDRLEN, buffer, size);
	sd = socket (PF_PACKET, SOCK_RAW, htons (ETH_P_ALL))) < 0);
	sendto (sd, rame, length...);
}
\end{lstlisting}

\subsection{Address translator}
\paragraph{}
The address translator is the subcomponent that assigns \gls{ipv4} and \gls{ipv6} to all devices in this system. For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns to them \gls{ipv4} addresses.
\paragraph{}
To do so, the translation algorithm is purely bijective. That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa. For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1. Once the packet is received and knowing it is not addressed to it. The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1. The \gls{ipv6} packet is created and reaches the destination which is the sensor node. Once the response is going back the scenario is the same.
\paragraph{}
The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address. When configuring the system, one should make sure that there is no two devices that have the same address real or virtual. Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system. The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.

\chapter{Findings}
\paragraph{}
The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system. Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services. The purpose of this chapter is to measure the performance of the system. It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
\section{Experiment}
\paragraph{}
To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system. \gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it. This means that there is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
\paragraph{Experiment 1:}
We will first have one mote in the \gls{wsn} that will be interacting with the outside world. We will start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the system and continue pinging the same mote. We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
\section{Data Gathering}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
\paragraph{Experiment 1}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes and 720 pings we done. Table \ref{table:exp1} shows the result of the experiment.

\begin{table}[htbp]
    \begin{tabular}{lll}
    \hline
    number of motes & Delay (ms) & Jitter (ms) \\\hline
    1               & 87.5         & 1.2         \\ 
    2               & 88.1         & 2.0         \\
    3               & 88.3        & 2.7         \\
    4               & 88.8         & 3.1         \\
    5               & 90.0         & 3.3         \\
    6               & 89.5         & 3.5         \\
    \end{tabular}
    \caption{Experiment 1 delay and jitter}
    \label{table:exp1}
\end{table}
For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row. The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval. The result is displayed in the table \ref{table:exp1}.

\paragraph{Experiment 2}
The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system. The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient. Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.

\begin{table}[htbp]
    \begin{tabular}{ll}
    \hline
    Delay (ms) & Jitter (ms) \\ \hline
    0.1        & 0.0         \\ 
    \end{tabular}
    \caption{Experiment 2 delay and jitter}
        \label{table:exp2}
\end{table}
The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nano seconds. Depending on the machine's load.
\section{Interpretation of Results}
The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
 nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by the 

@@

This project enabled us to establish an infrastructure as part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the smart. At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world which may be the utility for instance. This two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository which monitors the home's consumption and this device can be used for simulating a real world \gls{ami}. The other part of the communication is the ability of clients to control appliances status by turning them On and Off.
The second contribution of this work the smart management of the heterogeneity. The communication from the mote to the mobile client is built on a set of technologies and the journey of a simple \gls{tcp} datagram move through many different data link layer standards and different network technologies. Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame. Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame. Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that encapsulated into a Wi-Fi frame and sent until it reaches the mobile. That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
The \glsreset{wsn}\gls{wsn} was built making use of the very recent \glsreset{6lowpan}\gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices. The \gls{wsn} is autonomous in the sense that it can work independently and indefinitely by monitoring the home appliances. 
All in all, this project as a big picture is a testbed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middleware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users remote monitoring of their smart homes and control. This was possible by making the system as part of the \glsreset{iot}\gls{iot}.

@@

This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}. This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}. \gls{dr} is one of the most important topics in smart grids. It enables the utility and households to communicate with a set of services that any of them may request from the other. The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}. But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
Demand reponse integration is a future work for this study. \gls{dr} has a set of service that have to be implemented in the smart home packages. In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity. The consumer when in need of high priority in energy may ask the utility to provide such to provide a minimum of energy that will be guaranteed. Of course this means at a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems. Consumers can also help utilities at difficult times. For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to significantly cut their consumption and as a result be billed a reduced cost for the electricity that is consumed. In general, the smart home aims to reduce energy consumption and as the utility at times may need to reduce its energy production especially in peak hours, the \gls{dr} will help the utility ask the smart home to reduce the home consumption.
As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, there is a need to integrate \glsreset{rfid}\gls{rfid} within this system as a future work. The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification that will allow more flexibility to the system, as the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on the \gls{ipv6} which may change from time to time and therefore cannot keep track of the each device's information and monitor it reliably.
This system provides near real-time sensory data that is not much exploited intelligently. The only use of this data is to enable users to monitor their home consumption. As far as the system is concerned, there is a need for a more intelligent exploitation of this very rich data that can be used by intelligent algorithm to automatically and autonomously reduce the household's energy consumption. Such algorithms can be based on machine learning and data mining techniques making the direction of this a cross disciplinary project aiming to reduce the consumption of household without reducing the utility of users. In other words, build algorithms that will reduce consumption of households while the user is not noticing change.
Before building new services on top of this system, one should test different aspects of the system. In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance. As a future work, one should test different aspects of the system such as the reliability and availability of the system. More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
The system provides different services, but there is not authentication or authorization involved. This means that any users knowing the mote connected to the appliance he/she is interested in, he/she can control it and there is no existing party or component that will control such a communication. As a future work, the control of the system's access should be monitored in order for the system to be more secure and robust to external attacks.

\backmatter
\singlespacing
\addcontentsline{toc}{section}{References}
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}

\doublespacing
\appendix
\chapter{How to install the system}
\section{Hardware needed}
In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services. The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi. A smartphone or a separate computer with Android emulator may de used.

\section{Software installation}

\subsection{Operating system}
The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine using Cygwin or other tools
\subsection{TinyOs Installation}
\begin{enumerate}
\item Execute: sudo gedit /etc/apt/source.list
\item Add to the opened file: deb http://hinrg.cs.jhu.edu/tinyos hardy main
\item sudo apt-get update 
\item sudo apt-get install tinyos-2.1.1
\item sudo gedit ~/.bashrc
\item add the following line: \\
export TOSDIR=\$TOSROOT/tos \\
export CLASSPATH=\$TOSROOT/support/sdk/java/tinyos.jar:.\$CLASSPATH \\
export MAKERULES=\$TOSROOT/support/make/Makerules \\
export PATH=/opt/msp430/bin:\$PATH \\
source /opt/tinyos-2.1.1/tinyos.sh \\

\item sudo chmod 777 \$TOSROOT
\end{enumerate}

\subsection{Configuring \gls{blip}}
\glsreset{blip}\gls{blip} should be compiled before being used. To do so, follow these instructions:
\begin{enumerate}
\item cd \$TOSROOT/support/sdk/c/sf \\
./bootstrap \\
./configure \\
make 
\item cd \$TOSROOT/support/sdk/c/blip \\
./bootstrap.sh \\
./configure \\
make 
\end{enumerate}

\subsection{Install Software in Motes}
The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
\begin{enumerate}
\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
\item cd \$TOSROOT/apps/IPBaseStation
\item make micaz blip install  mib520, /dev/ttyUSB0
\end{enumerate}
The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
\begin{enumerate}
\item Make sure you have SVN by typing svn --version
\item If you do not have SVN type sudo apt-get install subversion
\item download the file:
\item cd
\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
\item cd grid-aui-tinyos//programming/TCPEcho
\end{enumerate}
Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:

\begin{enumerate}
\item  make micaz blip install.*NUM mib520,/dev/tty*USB
\item make clean
\end{enumerate}
Note that *NUM refers to the NODE\_ID. For the first mote choose 1, for the second choose 2...

\subsection{Install Packet Transformation process}
To install this component, you must do the following:
\begin{enumerate}
\item Make sure you have CMake, type sudo apt-get install cmake
\item cd
\item grid-aui-tinyos//programming/netfiltering
\item cmake CMakeLists.txt
\item make
\end{enumerate}

\subsection{Install Gateway Sensor Data TCP Server}
To install the server you must do the following:
\begin{enumerate}
\item cd
\item grid-aui-tinyos//programming/GatewayServer/src
\item make
\end{enumerate}

\chapter{Run the System}
There is a set of instructions to follow in order as some of the components must be alive only when others are already alive. To run the system, do the following:
\begin{enumerate}
\item To turn on the Tunnel do the following
\subitem cd \$TOSROOT/support/sdk/c/blip/ 
\subitem sudo driver/ip-driver /dev/ttyUSB0 micaz. \\
 You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the figure \ref{fig:tunnel_output}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.9]{img/tunnel_output.jpg}
\caption{IPv6 over USB tunnel execution output}
\label{fig:tunnel_output}
\end{figure}

\item To Run the Gateway Sensor Data TCP server do the following:
\subitem cd
\subitem cd grid-aui-tinyos/programming/GatewayServer/src
\subitem java Server

\item To Enable IP forwarding do the following:
\subitem sudo bash
\subitem echo 1 \textgreater /proc/sys/net/ipv4/ip\_forward
\subitem exit

\item  To drop all incoming packets: sudo iptables -P INPUT DROP
\item To disable ICPMP REDIRECT TO HOST:
\subitem sudo bash
\subitem /sbin/sysctl -w net.ipv4.conf.all.accept\_redirects=0
\subitem /sbin/sysctl -w net.ipv4.conf.all.send\_redirects=0
\subitem /sbin/sysctl -w net.ipv6.conf.all.accept\_redirects=0
\subitem /sbin/sysctl -w net.ipv6.conf.all.send\_redirects=0
\subitem exit

\item Run the Packet transformation process
\subitem cd
\subitem cd grid-aui-tinyos/programming/netfiltering
\subitem sudo my\_sniffer

\item Turn on the wireless sensor nodes one by one
\end{enumerate}

\paragraph{}
You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
